### 2015-09-30 final_submission/manuscript.tex
updated manuscript template

\documentclass[titlepage]{article}

\usepackage{amsmath}

\title{Effective Vaccination Strategies for an H1N1 Epidemic in Ithaca, New York}
\author{Prepared by Paul Chesnais (pmc85), Christopher Silvia (cps232), \\ and Ryan Vogan (rcv39) in completion of Project 01 for MATH 3610}
\date{September 30, 2015}
\begin{document}
\maketitle

\section{Introduction}

\section{Deterministic Model}
\subsection{Algorithm}
	//TODO
\subsection{Assumptions and Parameter Justifications}
	//TODO
\subsection{Predictions}
	//TODO
\subsection{Validation and Robustness Testing}
	//TODO
\subsection{Strengths and Weaknesses}
	//TODO
\subsection{Future Work}
	//TODO

\section{Stochastic Model}
\subsection{Algorithm}
	//TODO
\subsection{Assumptions and Parameter Justifications}
	//TODO
\subsection{Predictions}
	//TODO
\subsection{Validation and Robustness Testing}
	//TODO
\subsection{Strengths and Weaknesses}
	//TODO
\subsection{Future Work}
	//TODO

\section{Conclusions}
	//TODO
\section{References}
	//TODO
\section{Source Code}
	//TODO
\section{Individual Contributions}
	//TODO

\end{document}